---
title: "Model input (*lavaan* (R) and Mplus)"
output: 
  html_document:
    toc:  true
    toc_float: true
---

---

## Setup
Below you can find the model syntax for specifying the state-trait models, both in long- and wide-format, using *lavaan* (R) and Mplus. Here, you can also find model syntax for the additional analyses and simulations that were performed in the paper. For SEM models with and without stationarity constraints, the data must be specified in wide-format. For ML-SEM models, data must be specified in long-format. 

---

## CUTS model {.tabset .tabset-fade}

The tabs below contain *lavaan* and Mplus model syntax for the SEM CUTS model, the stationary SEM CUTS model, and the ML-SEM CUTS model.

### SEM CUTS

#### *lavaan*

```{r sem-cuts, eval = F}
sem_cuts <- '
  # Common trait factor loadings
  CT =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
        IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
        IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
        IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
        IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # Common trait variance
  CT ~~ CT
  
  # Unique traits factor loadings
  UT_IDEA =~ IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  UT_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  UT_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  UT_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  UT_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5
  
  # Unique trait variances
  UT_IDEA ~~ UT_IDEA
  UT_UITS ~~ UT_UITS
  UT_TEVR ~~ UT_TEVR
  UT_BELA ~~ UT_BELA
  UT_VERA ~~ UT_VERA
  
  # Common state factor loadings
  CS1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  CS2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  CS3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  CS4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  CS5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  CS1 ~~ CS1
  CS2 ~~ CS2
  CS3 ~~ CS3
  CS4 ~~ CS4
  CS5 ~~ CS5
  
  # Unique state variances (included by default)
  
  # Means (included by default through `meanstructure = T` argument)
'

fit_sem_cuts <- sem(model = sem_cuts, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_sem_cuts)
```

#### Mplus 

```
TITLE:      The SEM CUTS model.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Common trait factor loadings 
            CT BY IDEA1-VERA5; 

            ! Common trait variance (default)
            CT;
            
            ! Unique traits factor loadings
            UT_IDEA BY IDEA1@1 IDEA2 IDEA3 IDEA4 IDEA5; 
            UT_UITS BY UITS1@1 UITS2 UITS3 UITS4 UITS5;  
            UT_TEVR BY TEVR1@1 TEVR2 TEVR3 TEVR4 TEVR5; 
            UT_BELA BY BELA1@1 BELA2 BELA3 BELA4 BELA5;
            UT_VERA BY VERA1@1 VERA2 VERA3 VERA4 VERA5;
	        
            ! Unique trait variances (default)
            UT_IDEA-UT_VERA;
            
            ! Common state factor loadings
            CS1 BY IDEA1-VERA1;
            CS2 BY IDEA2-VERA2;
            CS3 BY IDEA3-VERA3;
            CS4 BY IDEA4-VERA4;
            CS5 BY IDEA5-VERA5;

            ! Common state variance (default)
            CS1-CS5;
            
            ! Unique state variances (default)
            IDEA1-VERA5;
    
            ! Means
            [IDEA1-VERA5];

OUTPUT:     SAMPSTAT TECH1 ;
```

---

### Stationary SEM CUTS

#### *lavaan*

```{r stationary-cuts, eval = F}
stationary_cuts <- '
  # Common trait factor loadings (constrained)
  CT =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
        g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
        g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
        g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
        g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # Common trait variance
  CT ~~ CT
  
  # Unique traits factor loadings (constrained)
  UT_IDEA =~ 1*IDEA_1 + 1*IDEA_2 + 1*IDEA_3 + 1*IDEA_4 + 1*IDEA_5
  UT_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  UT_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  UT_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  UT_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5
  
  # Unique trait variances
  UT_IDEA ~~ UT_IDEA
  UT_UITS ~~ UT_UITS
  UT_TEVR ~~ UT_TEVR
  UT_BELA ~~ UT_BELA
  UT_VERA ~~ UT_VERA
  
  # Common state factor loadings (constrained)
  CS1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  CS2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  CS3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  CS4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  CS5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  CS1 ~~ vcs*CS1
  CS2 ~~ vcs*CS2
  CS3 ~~ vcs*CS3
  CS4 ~~ vcs*CS4
  CS5 ~~ vcs*CS5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit_stationary_cuts <- sem(model = stationary_cuts, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_stationary_cuts)
```

#### Mplus 

```
TITLE:      The SEM CUTS model with stationarity constraints over time.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Common trait factor loadings (constrained)
            CT BY IDEA1@1 UITS1 TEVR1 BELA1 VERA1 (g1-g5); 
            CT BY IDEA2@1 UITS2 TEVR2 BELA2 VERA2 (g1-g5); 
            CT BY IDEA3@1 UITS3 TEVR3 BELA3 VERA3 (g1-g5);
            CT BY IDEA4@1 UITS4 TEVR4 BELA4 VERA4 (g1-g5);
            CT BY IDEA5@1 UITS5 TEVR5 BELA5 VERA5 (g1-g5);
            
            ! Common trait variance (default)
            CT;
            
            ! Unique trait factor loadings (constrained)
            UT_IDEA BY IDEA1@1 IDEA2@1 IDEA3@1 IDEA4@1 IDEA5@1; 
            UT_UITS BY UITS1@1 UITS2@1 UITS3@1 UITS4@1 UITS5@1;  
          	UT_TEVR BY TEVR1@1 TEVR2@1 TEVR3@1 TEVR4@1 TEVR5@1; 
          	UT_BELA BY BELA1@1 BELA2@1 BELA3@1 BELA4@1 BELA5@1;
          	UT_VERA BY VERA1@1 VERA2@1 VERA3@1 VERA4@1 VERA5@1;
          	
          	! Common state factor loadings (constrained) 
          	CS1 BY IDEA1 UITS1 TEVR1 BELA1 VERA1 (k1-k5);
            CS2 BY IDEA2 UITS2 TEVR2 BELA2 VERA2 (k1-k5);
            CS3 BY IDEA3 UITS3 TEVR3 BELA3 VERA3 (k1-k5);
            CS4 BY IDEA4 UITS4 TEVR4 BELA4 VERA4 (k1-k5);
            CS5 BY IDEA5 UITS5 TEVR5 BELA5 VERA5 (k1-k5);

            ! Common state variances (constrained)
            CS1-CS5 (vcs);
            
            ! Unique state variances (constrained) 
            IDEA1 UITS1 TEVR1 BELA1 VERA1 (vus1-vus5); 
            IDEA2 UITS2 TEVR2 BELA2 VERA2 (vus1-vus5);
            IDEA3 UITS3 TEVR3 BELA3 VERA3 (vus1-vus5);
            IDEA4 UITS4 TEVR4 BELA4 VERA4 (vus1-vus5);
            IDEA5 UITS5 TEVR5 BELA5 VERA5 (vus1-vus5);
    
            ! Means (constrained) 
            [IDEA1 UITS1 TEVR1 BELA1 VERA1] (m1-m5); 
            [IDEA2 UITS2 TEVR2 BELA2 VERA2] (m1-m5);
            [IDEA3 UITS3 TEVR3 BELA3 VERA3] (m1-m5);
            [IDEA4 UITS4 TEVR4 BELA4 VERA4] (m1-m5);
            [IDEA5 UITS5 TEVR5 BELA5 VERA5] (m1-m5);
            
OUTPUT:     SAMPSTAT TECH1;
```
---

### ML-SEM CUTS

#### *lavaan*

```{r mlsem-cuts, eval = F}
mlsem_cuts <- '
  level: 1
      cs =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      ct =~ IDEA + UITS + TEVR + BELA + VERA
'

fit_mlsem_cuts <- sem(model = mlsem_cuts, data = dat_LS_long, cluster = "id")
summary(fit_mlsem_cuts)
```

#### Mplus

```
TITLE:      The SEM CUTS model with stationarity constraints over time.

DATA:       FILE = LISS_LS_long.dat;
	
VARIABLE:   NAMES = id wave person house IDEA UITS TEVR BELA VERA;
            USEVARIABLES = IDEA-VERA;
            CLUSTER = id;
            MISSING = .;

ANALYSIS:   TYPE = TWOLEVEL;
            ESTIMATOR = ML;

MODEL:      %WITHIN%
            CS BY IDEA-VERA;
            
            %BETWEEN%
            CT BY IDEA-VERA;

OUTPUT:     TECH1;
```


---

## IT model {.tabset .tabset-fade}

The tabs below contain *lavaan* and Mplus model syntax for the SEM IT model, the stationary SEM IT model, and the ML-SEM IT model.

### SEM IT 

#### *lavaan*

```{r sem-it, eval = F}
sem_it <- '
  # Indicator-specific trait factor loadings
  IT_IDEA =~ IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  IT_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  IT_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  IT_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  IT_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5
  
  # Unique trait variances and covariances
  IT_IDEA ~~ IT_IDEA + IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_UITS ~~ IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_TEVR ~~ IT_TEVR + IT_BELA + IT_VERA
  IT_BELA ~~ IT_BELA + IT_VERA
  IT_VERA ~~ IT_VERA
  
  # Common state factor loadings
  CS1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  CS2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  CS3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  CS4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  CS5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  CS1 ~~ CS1
  CS2 ~~ CS2
  CS3 ~~ CS3
  CS4 ~~ CS4
  CS5 ~~ CS5
  
  # Unique state variances (included by default)
  
  # Means (included by default through `meanstructure = T` argument)
'

fit_sem_it <- sem(model = sem_it, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_sem_it)
```

#### Mplus


```
TITLE:      The SEM IT model.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Indicator-specific trait factor loadings
          	IT1 BY IDEA1@1 IDEA2 IDEA3 IDEA4 IDEA5;
          	IT2 BY UITS1@1 UITS2 UITS3 UITS4 UITS5;
          	IT3 BY TEVR1@1 TEVR2 TEVR3 TEVR4 TEVR5;
          	IT4 BY BELA1@1 BELA2 BELA3 BELA4 BELA5;
          	IT5 BY VERA1@1 VERA2 VERA3 VERA4 VERA5;
          	
          	! Correlations between indicator-specific trait factors
          	IT1 WITH IT2-IT5;
          	IT2 WITH IT3-IT5;
          	IT3 WITH IT4 IT5;
          	IT4 WITH IT5;
          	
          	! Common state factor loadings
          	CS1 BY IDEA1 UITS1 TEVR1 BELA1 VERA1;
          	CS2 BY IDEA2 UITS2 TEVR2 BELA2 VERA2;
          	CS3 BY IDEA3 UITS3 TEVR3 BELA3 VERA3;
          	CS4 BY IDEA4 UITS4 TEVR4 BELA4 VERA4;
            CS5 BY IDEA5 UITS5 TEVR5 BELA5 VERA5;
            
            ! Unique states variances (default)
            IDEA1-VERA5;

OUTPUT:     SAMPSTAT  TECH1;

```

---

### Stationary SEM IT

#### *lavaan*

```{r stationary-it, eval = F}
stationary_it <- '
  # Indicator-specific trait factor loadings
  IT_IDEA =~ 1*IDEA_1 + 1*IDEA_2 + 1*IDEA_3 + 1*IDEA_4 + 1*IDEA_5
  IT_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  IT_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  IT_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  IT_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5
  
  # Unique trait variances and covariances
  IT_IDEA ~~ IT_IDEA + IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_UITS ~~ IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_TEVR ~~ IT_TEVR + IT_BELA + IT_VERA
  IT_BELA ~~ IT_BELA + IT_VERA
  IT_VERA ~~ IT_VERA
  
  # Common state factor loadings (constrained)
  CS1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  CS2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  CS3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  CS4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  CS5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  CS1 ~~ vcs*CS1
  CS2 ~~ vcs*CS2
  CS3 ~~ vcs*CS3
  CS4 ~~ vcs*CS4
  CS5 ~~ vcs*CS5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit_stationary_it <- sem(model = stationary_it, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_stationary_it)
```

#### Mplus 

```
TITLE:      The SEM IT model with stationarity constraints over time.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Indicator-specific trait factor loadings
            IT1 BY IDEA1@1 IDEA2@1 IDEA3@1 IDEA4@1 IDEA5@1;
            IT2 BY UITS1@1 UITS2@1 UITS3@1 UITS4@1 UITS5@1;
            IT3 BY TEVR1@1 TEVR2@1 TEVR3@1 TEVR4@1 TEVR5@1;
            IT4 BY BELA1@1 BELA2@1 BELA3@1 BELA4@1 BELA5@1;
	          IT5 BY VERA1@1 VERA2@1 VERA3@1 VERA4@1 VERA5@1;
	
	          ! Correlations between indicator-specific trait factors
	          IT1 WITH IT2-IT5;
	          IT2 WITH IT3-IT5;
            IT3 WITH IT4 IT5;
	          IT4 WITH IT5;
	          
	          ! Common state factor loadings (constrained)
	          CS1 BY IDEA1 UITS1 TEVR1 BELA1 VERA1 (k1-k5);
	          CS2 BY IDEA2 UITS2 TEVR2 BELA2 VERA2 (k1-k5);
	          CS3 BY IDEA3 UITS3 TEVR3 BELA3 VERA3 (k1-k5);
	          CS4 BY IDEA4 UITS4 TEVR4 BELA4 VERA4 (k1-k5);
            CS5 BY IDEA5 UITS5 TEVR5 BELA5 VERA5 (k1-k5);
	
            ! Common states variances (constrained)
            CS1-CS5 (vcs);
            
            ! Unique states factor loadings (constrained)
            IDEA1 UITS1 TEVR1 BELA1 VERA1 (vus1-vus5); 
            IDEA2 UITS2 TEVR2 BELA2 VERA2 (vus1-vus5);
            IDEA3 UITS3 TEVR3 BELA3 VERA3 (vus1-vus5);
            IDEA4 UITS4 TEVR4 BELA4 VERA4 (vus1-vus5);
            IDEA5 UITS5 TEVR5 BELA5 VERA5 (vus1-vus5);
    
            ! Means (constrained) 
            [IDEA1 UITS1 TEVR1 BELA1 VERA1] (m1-m5); 
            [IDEA2 UITS2 TEVR2 BELA2 VERA2] (m1-m5);
            [IDEA3 UITS3 TEVR3 BELA3 VERA3] (m1-m5);
            [IDEA4 UITS4 TEVR4 BELA4 VERA4] (m1-m5);
            [IDEA5 UITS5 TEVR5 BELA5 VERA5] (m1-m5);

OUTPUT:     SAMPSTAT TECH1; 
```

---

### ML-SEM IT 

#### *lavaan*

```{r mlsem-it, eval = F}
mlsem_it <- '
  level: 1
      cs =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      IDEA ~~ IDEA + UITS + TEVR + BELA + VERA
      UITS ~~ UITS + TEVR + BELA + VERA
      TEVR ~~ TEVR + BELA + VERA
      BELA ~~ BELA + VERA
      VERA ~~ VERA
'

fit_mlsem_it <- sem(model = mlsem_it, data = dat_LS_long, cluster = "id")
summary(fit_mlsem_it)
```

#### Mplus

```
TITLE:      The ML-SEM IT model.

DATA:       FILE = LISS_LS_long.dat;
	
VARIABLE:   NAMES = id wave person house IDEA UITS TEVR BELA VERA;
            USEVARIABLES = IDEA-VERA;
            CLUSTER = id;
            MISSING = .;

ANALYSIS:   TYPE = TWOLEVEL;
            ESTIMATOR = ML;

MODEL:      %WITHIN%
            cs BY IDEA-VERA;
            
            %BETWEEN%
            IDEA-VERA; ! Indicator-specific trait factors
            
            ! Correlations between IT factors
          	IDEA WITH UITS-VERA;
          	UITS WITH TEVR-VERA;
          	TEVR WITH BELA VERA;
          	BELA WITH VERA;

OUTPUT:     SAMPSTAT TECH1;
```

---

## RI model{.tabset .tabset-fade}

The tabs below contain *lavaan* and Mplus model syntax for the SEM RI model, the stationary SEM RI model, and the ML-SEM RI model.

### SEM RI

#### *lavaan*

```{r sem-RI, eval = F}
sem_RI <- '
  # General factor loadings
  G =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
       IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
       IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
       IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
       IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # General factor variance
  G ~~ G
  
  # Trait Residual factor loadings
  # IDEA serves as the reference indicator
  TR_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  TR_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  TR_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  TR_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5

  # Trait Residual factor variances
  TR_UITS ~~ TR_UITS
  TR_TEVR ~~ TR_TEVR
  TR_BELA ~~ TR_BELA
  TR_VERA ~~ TR_VERA
  
  # Allow covariances between trait residual factors
  TR_UITS ~~ TR_TEVR + TR_BELA + TR_VERA
  TR_TEVR ~~ TR_BELA + TR_VERA
  TR_BELA ~~ TR_VERA
  
  # Occasion factor loadings
  O1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  O2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  O3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  O4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  O5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  O1 ~~ O1
  O2 ~~ O2
  O3 ~~ O3
  O4 ~~ O4
  O5 ~~ O5
  
  # Unique state variances (included by default)
  
  # Means (included by default through `meanstructure = T` argument)
'

fit_sem_RI <- sem(model = sem_RI, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_sem_RI)
```

#### Mplus 

```
TITLE:      The SEM RI model.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! General factor factor loading
            G BY IDEA1-VERA5;
            
            ! Trait residual factor loading
            ! IDEA serves as the reference indicator
            TR_UITS BY UITS1 UITS2 UITS3 UITS4 UITS5;
            TR_TEVR BY TEVR1 TEVR2 TEVR3 TEVR4 TEVR5;
            TR_BELA BY BELA1 BELA2 BELA3 BELA4 BELA5;
            TR_VERA BY VERA1 VERA2 VERA3 VERA4 VERA5;

            ! Allow covariances between trait residual factors
            TR_UITS WITH TR_TEVR TR_BELA TR_VERA;
            TR_TEVR WITH TR_BELA TR_VERA;
            TR_BELA WITH TR_VERA;
	
	          ! Occasion factors
	          O1 BY IDEA1-VERA1;
	          O2 BY IDEA2-VERA2;
	          O3 BY IDEA3-VERA3;
	          O4 BY IDEA4-VERA4;
	          O5 BY IDEA5-VERA5; 

OUTPUT:     SAMPSTAT TECH1;
```


---

### Stationary SEM RI

#### *lavaan*

```{r stationary-RI, eval = F}
stationary_RI <- '
  # General factor loadings
  G =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
       g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
       g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
       g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
       g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # General factor variance
  G ~~ G
  
  # Trait Residual factor loadings
  # IDEA serves as the reference indicator
  TR_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  TR_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  TR_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  TR_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5

  # Trait Residual factor variances
  TR_UITS ~~ TR_UITS
  TR_TEVR ~~ TR_TEVR
  TR_BELA ~~ TR_BELA
  TR_VERA ~~ TR_VERA
  
  # Allow covariances between trait residual factors
  TR_UITS ~~ TR_TEVR + TR_BELA + TR_VERA
  TR_TEVR ~~ TR_BELA + TR_VERA
  TR_BELA ~~ TR_VERA
  
  # Common state factor loadings (constrained)
  O1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  O2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  O3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  O4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  O5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  O1 ~~ vcs*O1
  O2 ~~ vcs*O2
  O3 ~~ vcs*O3
  O4 ~~ vcs*O4
  O5 ~~ vcs*O5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit_stationary_RI <- sem(model = stationary_RI, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_stationary_RI)
```

#### Mplus

```
TITLE:      The SEM RI model with stationary constraints over time.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! General factor (constrained)
            G BY IDEA1@1 UITS1 TEVR1 BELA1 VERA1 (l1-l5);
            G BY IDEA2@1 UITS2 TEVR2 BELA2 VERA2 (l1-l5);
            G BY IDEA3@1 UITS3 TEVR3 BELA3 VERA3 (l1-l5);
            G BY IDEA4@1 UITS4 TEVR4 BELA4 VERA4 (l1-l5);
            G BY IDEA5@1 UITS5 TEVR5 BELA5 VERA5 (l1-l5);
            
            ! Trait residual factors (constrained)
            ! IDEA serves as the reference indicator
            TR_UITS BY UITS1@1 UITS2@1 UITS3@1 UITS4@1 UITS5@1;
            TR_TEVR BY TEVR1@1 TEVR2@1 TEVR3@1 TEVR4@1 TEVR5@1;
            TR_BELA BY BELA1@1 BELA2@1 BELA3@1 BELA4@1 BELA5@1;
            TR_VERA BY VERA1@1 VERA2@1 VERA3@1 VERA4@1 VERA5@1;
            
            ! Covariances between method factors
            TR_UITS WITH TR_TEVR TR_BELA TR_VERA;
            TR_TEVR WITH TR_BELA TR_VERA;
            TR_BELA WITH TR_VERA;
            
            ! Occasion factors (constrained)
            O1 BY IDEA1@1 UITS1 TEVR1 BELA1 VERA1 (k1-k5);
            O2 BY IDEA2@1 UITS2 TEVR2 BELA2 VERA2 (k1-k5);
            O3 BY IDEA3@1 UITS3 TEVR3 BELA3 VERA3 (k1-k5);
            O4 BY IDEA4@1 UITS4 TEVR4 BELA4 VERA4 (k1-k5);
            O5 BY IDEA5@1 UITS5 TEVR5 BELA5 VERA5 (k1-k5);
            
            ! Occasion factor variances (constrained)
            O1-O5 (omega);
            
            ! Measurement error variances (constrained)
            IDEA1 UITS1 TEVR1 BELA1 VERA1 (sigma1-sigma5);
            IDEA2 UITS2 TEVR2 BELA2 VERA2 (sigma1-sigma5);
            IDEA3 UITS3 TEVR3 BELA3 VERA3 (sigma1-sigma5);
            IDEA4 UITS4 TEVR4 BELA4 VERA4 (sigma1-sigma5);
            IDEA5 UITS5 TEVR5 BELA5 VERA5 (sigma1-sigma5);
            
            ! Means (constrained)
            [IDEA1 UITS1 TEVR1 BELA1 VERA1] (m1-m5);
            [IDEA2 UITS2 TEVR2 BELA2 VERA2] (m1-m5);
            [IDEA3 UITS3 TEVR3 BELA3 VERA3] (m1-m5);
            [IDEA4 UITS4 TEVR4 BELA4 VERA4] (m1-m5);
            [IDEA5 UITS5 TEVR5 BELA5 VERA5] (m1-m5);
            
OUTPUT:     SAMPSTAT TECH1;
```

---

### ML-SEM RI 

#### *lavaan*

```{r mlsem-RI, eval = F}
mlsem_RI <- '
  level: 1
      O =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      G =~ IDEA + UITS + TEVR + BELA + VERA
      
      IDEA ~~ 0*IDEA
      UITS ~~ UITS + TEVR + BELA + VERA
      TEVR ~~ TEVR + BELA + VERA
      BELA ~~ BELA + VERA
      VERA ~~ VERA
'

fit_mlsem_RI <- sem(model = mlsem_RI, data = dat_LS_long, cluster = "id")
summary(fit_mlsem_RI)
```

#### Mplus

```
TITLE:      The ML-SEM RI model.

DATA:       FILE = LISS_LS_long.dat;
	
VARIABLE:   NAMES = id wave person house IDEA UITS TEVR BELA VERA;
            USEVARIABLES = IDEA-VERA;
            CLUSTER = id;
            MISSING = .;

ANALYSIS:   TYPE = TWOLEVEL;
            ESTIMATOR = ML;

MODEL:	    %WITHIN%
            O BY IDEA-VERA;
            
            %BETWEEN%
            G BY IDEA-VERA;
            
            IDEA@0; ! IDEA serves as reference indicator
            UITS-VERA; ! Include trait residual factors
            UITS WITH TEVR-VERA; ! Include covariance between method residual factors
            TEVR WITH BELA VERA;
            BELA WITH VERA;
	
OUTPUT:     SAMPSTAT TECH1;
```


---

## CU model {.tabset .tabset-fade}

The tabs below contain *lavaan* and Mplus model syntax for the SEM CU model, and the stationary SEM CU model. Within the long-format ML-SEM framework, it is not possible to specify correlations between residuals over time. Therefore, an ML-SEM CU model is not identified. 

### SEM CU

#### *lavaan*

```{r sem-cu, eval = F}
sem_cu <- '
  # Trait factor loadings
  T =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
       IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
       IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
       IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
       IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # Trait factor variance
  T ~~ T
  
  # State residual factor loadings
  SR1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  SR2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  SR3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  SR4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  SR5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  SR1 ~~ SR1
  SR2 ~~ SR2
  SR3 ~~ SR3
  SR4 ~~ SR4
  SR5 ~~ SR5
  
  # Correlated measurement errors between the same methods over time 
  IDEA_1 ~~ IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  IDEA_2 ~~ IDEA_3 + IDEA_4 + IDEA_5
  IDEA_3 ~~ IDEA_4 + IDEA_5
  IDEA_4 ~~ IDEA_5
  
  UITS_1 ~~ UITS_2 + UITS_3 + UITS_4 + UITS_5  
  UITS_2 ~~ UITS_3 + UITS_4 + UITS_5 
  UITS_3 ~~ UITS_4 + UITS_5 
  UITS_4 ~~ UITS_5 
  
  TEVR_1 ~~ TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  TEVR_2 ~~ TEVR_3 + TEVR_4 + TEVR_5
  TEVR_3 ~~ TEVR_4 + TEVR_5
  TEVR_4 ~~ TEVR_5
  
  BELA_1 ~~ BELA_2 + BELA_3 + BELA_4 + BELA_5
  BELA_2 ~~ BELA_3 + BELA_4 + BELA_5
  BELA_3 ~~ BELA_4 + BELA_5
  BELA_4 ~~ BELA_5
  
  VERA_1 ~~ VERA_2 + VERA_3 + VERA_4 + VERA_5
  VERA_2 ~~ VERA_3 + VERA_4 + VERA_5
  VERA_3 ~~ VERA_4 + VERA_5
  VERA_4 ~~ VERA_5
'

fit_sem_cu <- sem(model = sem_cu, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_sem_cu)
```

#### Mplus 

```
TITLE:      The SEM CU model.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Trait factor loadings
            T BY IDEA1-VERA5;
            
            ! Trait variance (default)
            T;
            
            ! State Residual factor loadings
            SR1 BY IDEA1-VERA1;
            SR2 BY IDEA2-VERA2;
            SR3 BY IDEA3-VERA3;
            SR4 BY IDEA4-VERA4;
            SR5 BY IDEA5-VERA5;
            
            ! State Residual variances (default)
            SR1-SR5;
            
            ! Correlated measurement errors between the same methods over time
            IDEA1 WITH IDEA2 IDEA3 IDEA4 IDEA5; ! Method 1
            IDEA2 WITH IDEA3 IDEA4 IDEA5;
            IDEA3 WITH IDEA4 IDEA5;
            IDEA4 WITH IDEA5;

            UITS1 WITH UITS2 UITS3 UITS4 UITS5;
            UITS2 WITH UITS3 UITS4 UITS5;
            UITS3 WITH UITS4 UITS5;
            UITS4 WITH UITS5;

            TEVR1 WITH TEVR2 TEVR3 TEVR4 TEVR5;
            TEVR2 WITH TEVR3 TEVR4 TEVR5;
            TEVR3 WITH TEVR4 TEVR5;
            TEVR4 WITH TEVR5;
    
            BELA1 WITH BELA2 BELA3 BELA4 BELA5;
            BELA2 WITH BELA3 BELA4 BELA5;
            BELA3 WITH BELA4 BELA5;
            BELA4 WITH BELA5;

            VERA1 WITH VERA2 VERA3 VERA4 VERA5;
            VERA2 WITH VERA3 VERA4 VERA5;
            VERA3 WITH VERA4 VERA5;
            VERA4 WITH VERA5;

OUTPUT:     SAMPSTAT  TECH1; 
```

---

### Stationary SEM CU

#### *lavaan* 

```{r stationary-cu, eval = F}
stationary_cu <- '
  # Trait factor loadings
  T =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
       g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
       g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
       g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
       g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # Trait factor variance
  T ~~ T
  
  # State Residual factor loadings (constrained)
  SR1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  SR2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  SR3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  SR4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  SR5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # State Residual variances (constrained)
  SR1 ~~ vcs*SR1
  SR2 ~~ vcs*SR2
  SR3 ~~ vcs*SR3
  SR4 ~~ vcs*SR4
  SR5 ~~ vcs*SR5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5
  
  # Correlated measurement errors between the same methods over time 
  IDEA_1 ~~ cov1*IDEA_2 + cov1*IDEA_3 + cov1*IDEA_4 + cov1*IDEA_5
  IDEA_2 ~~ cov1*IDEA_3 + cov1*IDEA_4 + cov1*IDEA_5
  IDEA_3 ~~ cov1*IDEA_4 + cov1*IDEA_5
  IDEA_4 ~~ cov1*IDEA_5
  
  UITS_1 ~~ cov2*UITS_2 + cov2*UITS_3 + cov2*UITS_4 + cov2*UITS_5  
  UITS_2 ~~ cov2*UITS_3 + cov2*UITS_4 + cov2*UITS_5 
  UITS_3 ~~ cov2*UITS_4 + cov2*UITS_5 
  UITS_4 ~~ cov2*UITS_5 
  
  TEVR_1 ~~ cov3*TEVR_2 + cov3*TEVR_3 + cov3*TEVR_4 + cov3*TEVR_5 
  TEVR_2 ~~ cov3*TEVR_3 + cov3*TEVR_4 + cov3*TEVR_5
  TEVR_3 ~~ cov3*TEVR_4 + cov3*TEVR_5
  TEVR_4 ~~ cov3*TEVR_5
  
  BELA_1 ~~ cov4*BELA_2 + cov4*BELA_3 + cov4*BELA_4 + cov4*BELA_5
  BELA_2 ~~ cov4*BELA_3 + cov4*BELA_4 + cov4*BELA_5
  BELA_3 ~~ cov4*BELA_4 + cov4*BELA_5
  BELA_4 ~~ cov4*BELA_5
  
  VERA_1 ~~ cov5*VERA_2 + cov5*VERA_3 + cov5*VERA_4 + cov5*VERA_5
  VERA_2 ~~ cov5*VERA_3 + cov5*VERA_4 + cov5*VERA_5
  VERA_3 ~~ cov5*VERA_4 + cov5*VERA_5
  VERA_4 ~~ cov5*VERA_5
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit_stationary_cu <- sem(model = stationary_cu, data = dat_LS, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit_stationary_cu)
```

#### Mplus

```
TITLE:      The SEM CU model with stationarity constraints over time.
          
DATA:       FILE = LISS_LS_wide.dat;
	
VARIABLE:   NAMES = person house 
                    IDEA1 UITS1 TEVR1 BELA1 VERA1 
                    IDEA2 UITS2 TEVR2 BELA2 VERA2 
                    IDEA3 UITS3 TEVR3 BELA3 VERA3 
                    IDEA4 UITS4 TEVR4 BELA4 VERA4
                    IDEA5 UITS5 TEVR5 BELA5 VERA5; 
            MISSING = .;
            USEVARIABLES = IDEA1-VERA5;

ANALYSIS:   MODEL = NOCOV;

MODEL:      ! Trait factor loadings (constrained)
            T BY IDEA1@1 UITS1 TEVR1 BELA1 VERA1 (g1-g5);
            T BY IDEA2@1 UITS2 TEVR2 BELA2 VERA2 (g1-g5);
            T BY IDEA3@1 UITS3 TEVR3 BELA3 VERA3 (g1-g5);
            T BY IDEA4@1 UITS4 TEVR4 BELA4 VERA4 (g1-g5);
            T BY IDEA5@1 UITS5 TEVR5 BELA5 VERA5 (g1-g5);
            
            ! Trait variance (default)
            T;
            
            ! State Residual factors (constrained)
            SR1 BY IDEA1-VERA1 (k1-k5);
            SR2 BY IDEA2-VERA2 (k1-k5);
            SR3 BY IDEA3-VERA3 (k1-k5);
            SR4 BY IDEA4-VERA4 (k1-k5);
            SR5 BY IDEA5-VERA5 (k1-k5);
            
            ! State Residual variances (constrained)
            SR1-SR5 (vcs);
            
            ! Measurement error variances (constrained)
            IDEA1 UITS1 TEVR1 BELA1 VERA1 (vus1-vus5);
            IDEA2 UITS2 TEVR2 BELA2 VERA2 (vus1-vus5);
            IDEA3 UITS3 TEVR3 BELA3 VERA3 (vus1-vus5);
            IDEA4 UITS4 TEVR4 BELA4 VERA4 (vus1-vus5);
            IDEA5 UITS5 TEVR5 BELA5 VERA5 (vus1-vus5);
            
            ! Correlated measurement errors between the same methods over time (constrained)
            IDEA1 WITH IDEA2 IDEA3 IDEA4 IDEA5 (cov1);
            IDEA2 WITH IDEA3 IDEA4 IDEA5 (cov1);
            IDEA3 WITH IDEA4 IDEA5 (cov1);
            IDEA4 WITH IDEA5 (cov1);
            
            UITS1 WITH UITS2 UITS3 UITS4 UITS5 (cov2);
            UITS2 WITH UITS3 UITS4 UITS5 (cov2);
            UITS3 WITH UITS4 UITS5 (cov2);
            UITS4 WITH UITS5 (cov2);
            
            TEVR1 WITH TEVR2 TEVR3 TEVR4 TEVR5 (cov3);
            TEVR2 WITH TEVR3 TEVR4 TEVR5 (cov3);
            TEVR3 WITH TEVR4 TEVR5 (cov3);
            TEVR4 WITH TEVR5 (cov3);
    
   	        BELA1 WITH BELA2 BELA3 BELA4 BELA5 (cov4);
   	        BELA2 WITH BELA3 BELA4 BELA5 (cov4);
            BELA3 WITH BELA4 BELA5 (cov4);
            BELA4 WITH BELA5 (cov4);
            
            VERA1 WITH VERA2 VERA3 VERA4 VERA5 (cov5);
            VERA2 WITH VERA3 VERA4 VERA5 (cov5);
            VERA3 WITH VERA4 VERA5 (cov5);
            VERA4 WITH VERA5 (cov5);
            
            ! Means (constrained)
            [IDEA1 UITS1 TEVR1 BELA1 VERA1] (m1-m5);
            [IDEA2 UITS2 TEVR2 BELA2 VERA2] (m1-m5);
            [IDEA3 UITS3 TEVR3 BELA3 VERA3] (m1-m5);
            [IDEA4 UITS4 TEVR4 BELA4 VERA4] (m1-m5);
            [IDEA5 UITS5 TEVR5 BELA5 VERA5] (m1-m5);

OUTPUT:     SAMPSTAT TECH1;
```

## Simulations: Skewness {.tabset .tabset-fade}

To explore the cause of the minor discrepancies in the variance decomposition of the CUTS and CU models on the one hand, and the IT and RI models on the other, two datasets were generated using Mplus according to the ML-SEM CUTS model: The first dataset is multivariate normally distributed, while the data in the second dataset is positively skewed. Next, both datasets were analyzed using the ML-SEM CUTS and the ML-SEM IT models, and parameter estimates were compared. 

### Simulate: Skewed - Model: ML-SEM CUTS

The below model syntax generates a positively skewed dataset from the ML-SEM CUTS model, and analyses it subsequently using the ML-SEM CUTS model. 

```
TITLE: 
	Simulation - CUTS model, parameter estimates based on analyses on 
		Life Satisfaction survey from the LISS Panel
	
	Analysis - CUTS model, 5 indicators, waves 1-5, estimated in 
		long-format (implicit stationarity constraints)
	
	Optionally - MCAR or skewed distribution (through mixture options)

MONTECARLO:
    NAMES = y1-y5;
    NOBSERVATIONS = 21430;
    NREPS = 1;
    SEED = 20211027;
    NCSIZES = 1; ! Number of unique cluster sizes
    CSIZES = 4286 (5); ! Specify 4286 clusters of size 5
    ! PATMISS = y1(.1) y2(.15) y3(.2) y4(.25) y5(.3); ! Proportion missing for each variable
    ! PATPROBS = 1; 
    GENCLASSES = c(2); ! Name categorical latent variables for generation (2 classes)
    CLASSES = c(1); ! Categorical latent variable for estimation (1 class)
    REPSAVE = ALL; 
    SAVE = simulated_CUTS_long_free*.dat;

MODEL POPULATION:
    %WITHIN%
    %OVERALL% ! At within-level, only a single class exists 
	
    [c#1*0]; ! Threshold latent categorical variable
	
    ! Common state factor loadings
    cs BY y1*1 y2*1.055 y3*0.939 y4*0.749 y5*0.659; 
    
    ! Common state factor variance
    cs*0.351;

    ! Unique state factor variances
    y1*0.278;
    y2*0.287;
    y3*0.243;
    y4*0.512;
    y5*0.921;
	
    %BETWEEN%
    %OVERALL% 
    
    c#1*1;
    
    ! Common trait factor loadings
    ct BY y1*1 y2*0.989 y3*0.881 y4*0.893 y5*1.036; 

    ! Common trait factor variance and mean
    ct*0.848; 
    [ct*0];

    ! Unique trait factor variances
    y1*0.059;
    y2*0.103; 
    y3*0.075; 
    y4*0.235;
    y5*0.635; 

    ! Means
    [y1*5.014 y2*5.140 y3*5.459 y4*5.211 y5*4.496];
    
    ! Threshold latent categorical variable
    [c#1*-1]; ! Determines class size

    %c#1% ! Mixture analysis (impact skewed data)
    [ct*3]; ! Common trait mean increased to 3


ANALYSIS:
	MODEL = NOCOV; ! Overrule defaults
	TYPE = TWOLEVEL MIXTURE;
	ESTIMATOR = ML;

MODEL:
	%WITHIN%
    %OVERALL%
	cs BY y1-y5; 
	
	%BETWEEN%
    %OVERALL%
	ct BY y1-y5;

OUTPUT:
	TECH1 TECH8 TECH9;
```

### Data: Skewed - Model: ML-SEM IT

The below model syntax analyses the generated positively skewed dataset using the ML-SEM IT model. 

```
TITLE: 
	Data - Simulated (with skewness) from CUTS model
	
	Analysis - IT model, 5 indicators, waves 1-5, 
		Estimated in long-format (implicit stationarity constraints)

DATA: 
    FILE = simulated_CUTS_long_free1.dat;

VARIABLE:
	NAMES = y1-y5 class id;
	USEVARIABLES = y1-y5;
	CLUSTER = id;
    MISSING = ALL(999); 

ANALYSIS:
	TYPE = TWOLEVEL;
	ESTIMATOR = ML;

MODEL:
	%WITHIN%
	cs BY y1-y5; 
	
	%BETWEEN%
	y1-y5; ! Indicator-specific trait factor variances
	
	! Correlations between IT factors
	y1 WITH y2-y5;
	y2 WITH y3-y5;
	y3 WITH y4 y5;
	y4 WITH y5;

OUTPUT:
	TECH1 STDYX;
```

---

## Simulations: Missing {.tabset .tabset-fade}

To explore the cause of the minor discrepancies in the variance decomposition of the CUTS and CU models on the one hand, and the IT and RI models on the other, two datasets were generated using Mplus according to the ML-SEM CUTS model: The first dataset contains missing data (MCAR), while the second dataset is complete. Next, both datasets were analyzed using the ML-SEM CUTS and the ML-SEM IT models, and parameter estimates were compared. 

### Simulate: MCAR - Model: ML-SEM CUTS

The below model syntax generates a dataset from the ML-SEM CUTS model, with 10% to 30% missing data (MCAR) for each variable. It subsequently analyses the generated data using the ML-SEM CUTS model. 

```
TITLE: 
	Simulation - CUTS model, parameter estimates based on analyses on 
		Life Satisfaction survey from the LISS Panel
	
	Analysis - CUTS model, 5 indicators, waves 1-5, estimated in 
		long-format (implicit stationarity constraints)
	
	Optionally - MCAR or skewed distribution (through mixture options)

MONTECARLO:
    NAMES = y1-y5;
    NOBSERVATIONS = 21430;
    NREPS = 1;
    SEED = 20211027;
    NCSIZES = 1; ! Number of unique cluster sizes
    CSIZES = 4286 (5); ! Specify 4286 clusters of size 5
    PATMISS = y1(.1) y2(.15) y3(.2) y4(.25) y5(.3); ! Proportion missing for each variable
    PATPROBS = 1; 
    REPSAVE = ALL; 
    SAVE = simulated_CUTS_long_free_MCAR*.dat;

MODEL POPULATION:
    %WITHIN%
    ! Common state factor loadings
    cs BY y1*1 y2*1.055 y3*0.939 y4*0.749 y5*0.659; 
    
    ! Common state factor variance
    cs*0.351;

    ! Unique state factor variances
    y1*0.278;
    y2*0.287;
    y3*0.243;
    y4*0.512;
    y5*0.921;
	
    %BETWEEN%
    ! Common trait factor loadings
    ct BY y1*1 y2*0.989 y3*0.881 y4*0.893 y5*1.036; 

    ! Common trait factor variance and mean
    ct*0.848; 
    [ct*0];

    ! Unique trait factor variances
    y1*0.059;
    y2*0.103; 
    y3*0.075; 
    y4*0.235;
    y5*0.635; 

    ! Means
    [y1*5.014 y2*5.140 y3*5.459 y4*5.211 y5*4.496];

ANALYSIS:
	MODEL = NOCOV; ! Overrule defaults
	TYPE = TWOLEVEL;
	ESTIMATOR = ML;

MODEL:
	%WITHIN%
	cs BY y1-y5; 
	
	%BETWEEN%
	ct BY y1-y5;

OUTPUT:
	TECH1 TECH8 TECH9;
```

### Data: MCAR - Model: ML-SEM IT

```
TITLE: 
	Data - Simulated (with MCAR) from CUTS model
	
	Analysis - IT model, 5 indicators, waves 1-5, 
		Estimated in long-format (implicit stationarity constraints)

DATA: 
    FILE = simulated_CUTS_long_free_MCAR1.dat;

VARIABLE:
	NAMES = y1-y5 class id;
	USEVARIABLES = y1-y5;
	CLUSTER = id;
    MISSING = ALL(999); 

ANALYSIS:
	TYPE = TWOLEVEL;
	ESTIMATOR = ML;

MODEL:
	%WITHIN%
	cs BY y1-y5; 
	
	%BETWEEN%
	y1-y5; ! Indicator-specific trait factor variances
	
	! Correlations between IT factors
	y1 WITH y2-y5;
	y2 WITH y3-y5;
	y3 WITH y4 y5;
	y4 WITH y5;

OUTPUT:
	TECH1 STDYX;
```

---

## Simulations: Convergence {.tabset .tabset-fade}

To explore potential convergence issues in the wide-format conceptualization of the CUTS model, we applied it to three other scales that have been included in the LISS panel data, namely the Positive and Negative Affect Scale (PANAS, 2 factors, 10 indicators per factor, 5 repeated measures, and a sample size of 4293), a self-esteem scale (SE, 1 factor, 10 indicators, 5 repeated measures, and a sample size of 4276), and a shortened self-esteem scale (SERU, 1 factor, 3 indicators, 5 repeated measures, and a sample size of 4276) by the Radboud University Nijmegen, the Netherlands. Below, you can find the Mplus model syntax for each analysis. 

### PANAS

```
TITLE: 
	The Common and Unique Trait State (CUTS) model
	PANAS survey from the LISS Panel
	20 indicators, 2 factors, wave 1, 2, 4, 6, 7
	Estimated in wide-format (free)

DATA:
	FILE = LISS_PANAS_wide.dat;
	TYPE = INDIVIDUAL; ! Wide-format (default)

	
VARIABLE:
	NAMES = id 
		    BELA1 VERO1 OPGE1 OVER1 STER1 SCHU1 PANI1
	        VIJA1 ENTH1 TROT1 PRIK1 ALER1 BESC1 GEIN1
	        NERV1 VAST1 OPLE1 ZENU1 ACTI1 BANG1
	         
	        BELA2 VERO2 OPGE2 OVER2 STER2 SCHU2 PANI2
	        VIJA2 ENTH2 TROT2 PRIK2 ALER2 BESC2 GEIN2
	        NERV2 VAST2 OPLE2 ZENU2 ACTI2 BANG2
	         
	        BELA3 VERO3 OPGE3 OVER3 STER3 SCHU3 PANI3
	        VIJA3 ENTH3 TROT3 PRIK3 ALER3 BESC3 GEIN3
	        NERV3 VAST3 OPLE3 ZENU3 ACTI3 BANG3
	         
	        BELA4 VERO4 OPGE4 OVER4 STER4 SCHU4 PANI4
	        VIJA4 ENTH4 TROT4 PRIK4 ALER4 BESC4 GEIN4
	        NERV4 VAST4 OPLE4 ZENU4 ACTI4 BANG4

            BELA5 VERO5 OPGE5 OVER5 STER5 SCHU5 PANI5
	        VIJA5 ENTH5 TROT5 PRIK5 ALER5 BESC5 GEIN5
	        NERV5 VAST5 OPLE5 ZENU5 ACTI5 BANG5;
	USEVARIABLES = BELA1-BANG5;
	MISSING = .;

ANALYSIS:
	MODEL = NOCOV;

MODEL:
	! Factor loadings common traits CT_PA and CT_NA
	CT_PA BY BELA1 BELA2 BELA3 BELA4 BELA5 
		     OPGE1 OPGE2 OPGE3 OPGE4 OPGE5 
		     STER1 STER2 STER3 STER4 STER5
		     ENTH1 ENTH2 ENTH3 ENTH4 ENTH5
		     TROT1 TROT2 TROT3 TROT4 TROT5
		     ALER1 ALER2 ALER3 ALER4 ALER5
		     GEIN1 GEIN2 GEIN3 GEIN4 GEIN5
		VAST1 VAST2 VAST3 VAST4 VAST5
		OPLE1 OPLE2 OPLE3 OPLE4 OPLE5
		ACTI1 ACTI2 ACTI3 ACTI4 ACTI5;
		
	
	CT_NA BY VERO1 VERO2 VERO3 VERO4 VERO5
		OVER1 OVER2 OVER3 OVER4 OVER5
		SCHU1 SCHU2 SCHU3 SCHU4 SCHU5
		PANI1 PANI2 PANI3 PANI4 PANI5
		VIJA1 VIJA2 VIJA3 VIJA4 VIJA5
		PRIK1 PRIK2 PRIK3 PRIK4 PRIK5
		BESC1 BESC2 BESC3 BESC4 BESC5
		NERV1 NERV2 NERV3 NERV4 NERV5
		ZENU1 ZENU2 ZENU3 ZENU4 ZENU5
		BANG1 BANG2 BANG3 BANG4 BANG5;
		
	
	! Variance CT (default)
	CT_PA; 
	CT_NA;
	
	! Factor loadings unique traits UT
	UT_BELA BY BELA1 BELA2 BELA3 BELA4 BELA5;
	UT_OPGE BY OPGE1 OPGE2 OPGE3 OPGE4 OPGE5;
	UT_STER BY STER1 STER2 STER3 STER4 STER5;
	UT_ENTH BY ENTH1 ENTH2 ENTH3 ENTH4 ENTH5;
	UT_TROT BY TROT1 TROT2 TROT3 TROT4 TROT5;
	UT_ALER BY ALER1 ALER2 ALER3 ALER4 ALER5;
	UT_GEIN BY GEIN1 GEIN2 GEIN3 GEIN4 GEIN5;
	UT_VAST BY VAST1 VAST2 VAST3 VAST4 VAST5;
	UT_OPLE BY OPLE1 OPLE2 OPLE3 OPLE4 OPLE5; 
	UT_ACTI BY ACTI1 ACTI2 ACTI3 ACTI4 ACTI5;
		
	UT_VERO BY VERO1 VERO2 VERO3 VERO4 VERO5;
	UT_OVER BY OVER1 OVER2 OVER3 OVER4 OVER5;
	UT_SCHU BY SCHU1 SCHU2 SCHU3 SCHU4 SCHU5;
	UT_PANI BY PANI1 PANI2 PANI3 PANI4 PANI5;
	UT_VIJA BY VIJA1 VIJA2 VIJA3 VIJA4 VIJA5;
	UT_PRIK BY PRIK1 PRIK2 PRIK3 PRIK4 PRIK5;
	UT_BESC BY BESC1 BESC2 BESC3 BESC4 BESC5;
	UT_NERV BY NERV1 NERV2 NERV3 NERV4 NERV5;
	UT_ZENU BY ZENU1 ZENU2 ZENU3 ZENU4 ZENU5;
	UT_BANG BY BANG1 BANG2 BANG3 BANG4 BANG5;

	! Factor loadings common state CS
	CS1_PA BY BELA1 OPGE1 STER1 ENTH1 TROT1 ALER1 GEIN1 VAST1 OPLE1 ACTI1;
	CS2_PA BY BELA2 OPGE2 STER2 ENTH2 TROT2 ALER2 GEIN2 VAST2 OPLE2 ACTI2;
	CS3_PA BY BELA3 OPGE3 STER2 ENTH3 TROT3 ALER3 GEIN3 VAST3 OPLE3 ACTI3;
	CS4_PA BY BELA4 OPGE4 STER4 ENTH4 TROT4 ALER4 GEIN4 VAST4 OPLE4 ACTI4;
    CS5_PA BY BELA5 OPGE5 STER5 ENTH5 TROT5 ALER5 GEIN5 VAST5 OPLE5 ACTI5;

	CS1_NA BY VERO1 OVER1 SCHU1 PANI1 VIJA1 PRIK1 BESC1 NERV1 ZENU1 BANG1;
	CS2_NA BY VERO2 OVER2 SCHU2 PANI2 VIJA2 PRIK2 BESC2 NERV2 ZENU2 BANG2;
	CS3_NA BY VERO3 OVER3 SCHU3 PANI3 VIJA3 PRIK3 BESC3 NERV3 ZENU3 BANG3;
    CS4_NA BY VERO4 OVER4 SCHU4 PANI4 VIJA4 PRIK4 BESC4 NERV4 ZENU4 BANG4;
    CS5_NA BY VERO5 OVER5 SCHU5 PANI5 VIJA5 PRIK5 BESC5 NERV5 ZENU5 BANG5;
	
	! Specify the unique states US 
	BELA1-BANG5;

OUTPUT:
	TECH1;
```

### SE

```
TITLE: 
	The Common and Unique Trait State (CUTS) model
	Rosenberg Self-Esteem scale from the LISS Panel
	10 indicators, wave 1, 2, 4, 6, 7
	Estimated in wide-format (free)

DATA:
	FILE = LISS_SE_wide2.dat;
	TYPE = INDIVIDUAL; ! Wide-format (default)

	
VARIABLE:
	NAMES = id 
		WAAR1 KWAL1 MISL1 NETZ1 TROT1 POSI1 TEVR1 RESP1 NUTT1 DEUG1 
		WAAR2 KWAL2 MISL2 NETZ2 TROT2 POSI2 TEVR2 RESP2 NUTT2 DEUG2 
		WAAR3 KWAL3 MISL3 NETZ3 TROT3 POSI3 TEVR3 RESP3 NUTT3 DEUG3 
		WAAR4 KWAL4 MISL4 NETZ4 TROT4 POSI4 TEVR4 RESP4 NUTT4 DEUG4
        WAAR5 KWAL5 MISL5 NETZ5 TROT5 POSI5 TEVR5 RESP5 NUTT5 DEUG5;
	USEVARIABLES = WAAR1-DEUG5;
	MISSING = .;

ANALYSIS:
	MODEL = NOCOV;
    ! STARTS = 20; 

MODEL:
	! Factor loadings common traits CT
	CT BY WAAR1 KWAL1 MISL1 NETZ1 TROT1 POSI1 TEVR1 RESP1 NUTT1 DEUG1 
	      WAAR2 KWAL2 MISL2 NETZ2 TROT2 POSI2 TEVR2 RESP2 NUTT2 DEUG2 
	      WAAR3 KWAL3 MISL3 NETZ3 TROT3 POSI3 TEVR3 RESP3 NUTT3 DEUG3 
	      WAAR4 KWAL4 MISL4 NETZ4 TROT4 POSI4 TEVR4 RESP4 NUTT4 DEUG4
          WAAR5 KWAL5 MISL5 NETZ5 TROT5 POSI5 TEVR5 RESP5 NUTT5 DEUG5;	
	
	! Variance CT (default)
	CT; 

	! Factor loadings unique traits UT
	UT_WAAR BY WAAR1 WAAR2 WAAR3 WAAR4 WAAR5;
	UT_KWAL BY KWAL1 KWAL2 KWAL3 KWAL4 KWAL5;
	UT_MISL BY MISL1 MISL2 MISL3 MISL4 MISL5;
	UT_NETZ BY NETZ1 NETZ2 NETZ3 NETZ4 NETZ5;
	UT_TROT BY TROT1 TROT2 TROT3 TROT4 TROT5;
	UT_POSI BY POSI1 POSI2 POSI3 POSI4 POSI5;
	UT_TEVR BY TEVR1 TEVR2 TEVR3 TEVR4 TEVR5; 
	UT_RESP BY RESP1 RESP2 RESP3 RESP4 RESP5;
	UT_NUTT BY NUTT1 NUTT2 NUTT3 NUTT4 NUTT5; 
	UT_DEUG BY DEUG1 DEUG2 DEUG3 DEUG4 DEUG5;
	
	! Factor loadings common state CS
	CS1 BY WAAR1 KWAL1 MISL1 NETZ1 TROT1 POSI1 TEVR1 RESP1 NUTT1 DEUG1; 
	CS2 BY WAAR2 KWAL2 MISL2 NETZ2 TROT2 POSI2 TEVR2 RESP2 NUTT2 DEUG2; 
	CS3 BY WAAR3 KWAL3 MISL3 NETZ3 TROT3 POSI3 TEVR3 RESP3 NUTT3 DEUG3; 
	CS4 BY WAAR4 KWAL4 MISL4 NETZ4 TROT4 POSI4 TEVR4 RESP4 NUTT4 DEUG4;
	CS5 BY WAAR5 KWAL5 MISL5 NETZ5 TROT5 POSI5 TEVR5 RESP5 NUTT5 DEUG5;

	! Specify the unique states US 
	WAAR1 KWAL1 MISL1 NETZ1 TROT1 POSI1 TEVR1 RESP1 NUTT1 DEUG1 
	WAAR2 KWAL2 MISL2 NETZ2 TROT2 POSI2 TEVR2 RESP2 NUTT2 DEUG2 
	WAAR3 KWAL3 MISL3 NETZ3 TROT3 POSI3 TEVR3 RESP3 NUTT3 DEUG3 
	WAAR4 KWAL4 MISL4 NETZ4 TROT4 POSI4 TEVR4 RESP4 NUTT4 DEUG4
    WAAR5 KWAL5 MISL5 NETZ5 TROT5 POSI5 TEVR5 RESP5 NUTT5 DEUG5;

OUTPUT:
	TECH1;
```

### SERU

```
TITLE: 
	The Common and Unique Trait State (CUTS) model
	Radboud University Self-Esteem scale from the LISS Panel
	3 indicators, wave 1, 2, 4, 6, 7
	Estimated in wide-format (free)

DATA:
	FILE = LISS_SERU_wide.dat;
	TYPE = INDIVIDUAL; ! Wide-format (default)

	
VARIABLE:
	NAMES = id 
		UITE1 GEVO1 CAPA1 
		UITE2 GEVO2 CAPA2 
		UITE3 GEVO3 CAPA3
     		UITE4 GEVO4 CAPA4 
     		UITE5 GEVO5 CAPA5; 
	USEVARIABLES = UITE1-CAPA5;
	MISSING = .;

ANALYSIS:
	MODEL = NOCOV;

MODEL:
	! Factor loadings common traits CT
	CT BY UITE1 GEVO1 CAPA1 
		  UITE2 GEVO2 CAPA2 
		  UITE3 GEVO3 CAPA3
     	  UITE4 GEVO4 CAPA4 
     	  UITE5 GEVO5 CAPA5; 
     		
	! Variance CT (default)
	CT; 

	! Factor loadings unique traits UT
	UT_UITE BY UITE1 UITE2 UITE3 UITE4 UITE5;
	UT_GEVO BY GEVO1 GEVO2 GEVO3 GEVO4 GEVO5;
	UT_CAPA BY CAPA1 CAPA2 CAPA3 CAPA4 CAPA5;
	
	! Factor loadings common state CS
	CS1 BY UITE1 GEVO1 CAPA1; 
	CS2 BY UITE2 GEVO2 CAPA2; 
	CS3 BY UITE3 GEVO3 CAPA3;
     	CS4 BY UITE4 GEVO4 CAPA4; 
     	CS5 BY UITE5 GEVO5 CAPA5; 
	
	! Specify the unique states US 
	UITE1-CAPA5; 

OUTPUT:
	TECH1;
```
