---
title: "Using lavaan (R)"
output: 
  html_document:
    toc:  true
    toc_float: true
    code_folding: show
---

---

## Setup
Below you can find code for specifying the state-trait models using *lavaan* in R. For SEM models with and without stationarity constraints, the data must be specified in wide-format. For ML-SEM models, data must be specified in long-format. 

---

## Common Unique Trait State model {.tabset .tabset-fade}

### SEM CUTS

```{r sem-cuts, eval = F}
sem.cuts <- '
  # Common trait factor loadings
  CT =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
        IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
        IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
        IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
        IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # Common trait variance
  CT ~~ CT
  
  # Unique traits factor loadings
  UT_IDEA =~ IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  UT_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  UT_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  UT_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  UT_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5
  
  # Unique trait variances
  UT_IDEA ~~ UT_IDEA
  UT_UITS ~~ UT_UITS
  UT_TEVR ~~ UT_TEVR
  UT_BELA ~~ UT_BELA
  UT_VERA ~~ UT_VERA
  
  # Common state factor loadings
  CS1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  CS2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  CS3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  CS4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  CS5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  CS1 ~~ CS1
  CS2 ~~ CS2
  CS3 ~~ CS3
  CS4 ~~ CS4
  CS5 ~~ CS5
  
  # Unique state variances (included by default)
  
  # Means (included by default through argument)
'

fit.sem.cuts <- sem(model = sem.cuts, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.sem.cuts)
```

---

### Stationary SEM CUTS

```{r stationary-cuts, eval = F}
stationary.cuts <- '
  # Common trait factor loadings (constrained)
  CT =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
        g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
        g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
        g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
        g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # Common trait variance
  CT ~~ CT
  
  # Unique traits factor loadings (constrained)
  UT_IDEA =~ 1*IDEA_1 + 1*IDEA_2 + 1*IDEA_3 + 1*IDEA_4 + 1*IDEA_5
  UT_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  UT_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  UT_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  UT_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5
  
  # Unique trait variances
  UT_IDEA ~~ UT_IDEA
  UT_UITS ~~ UT_UITS
  UT_TEVR ~~ UT_TEVR
  UT_BELA ~~ UT_BELA
  UT_VERA ~~ UT_VERA
  
  # Common state factor loadings (constrained)
  CS1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  CS2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  CS3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  CS4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  CS5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  CS1 ~~ vcs*CS1
  CS2 ~~ vcs*CS2
  CS3 ~~ vcs*CS3
  CS4 ~~ vcs*CS4
  CS5 ~~ vcs*CS5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit.stationary.cuts <- sem(model = stationary.cuts, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.stationary.cuts)
```

---

### ML-SEM CUTS
```{r mlsem-cuts, eval = F}
mlsem.cuts <- '
  level: 1
      cs =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      ct =~ IDEA + UITS + TEVR + BELA + VERA
'

fit.mlsem.cuts <- sem(model = mlsem.cuts, data = dat.liss.long, cluster = "id")
summary(fit.mlsem.cuts)
```

---


## Correlated Uniqueness model {.tabset .tabset-fade}


### SEM CU

```{r sem-cu, eval = F}
sem.cu <- '
  # Trait factor loadings
  T =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
       IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
       IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
       IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
       IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # Trait factor variance
  T ~~ T
  
  # State residual factor loadings
  SR1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  SR2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  SR3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  SR4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  SR5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  SR1 ~~ SR1
  SR2 ~~ SR2
  SR3 ~~ SR3
  SR4 ~~ SR4
  SR5 ~~ SR5
  
  # Correlated measurement errors between the same methods over time 
  IDEA_1 ~~ IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  IDEA_2 ~~ IDEA_3 + IDEA_4 + IDEA_5
  IDEA_3 ~~ IDEA_4 + IDEA_5
  IDEA_4 ~~ IDEA_5
  
  UITS_1 ~~ UITS_2 + UITS_3 + UITS_4 + UITS_5  
  UITS_2 ~~ UITS_3 + UITS_4 + UITS_5 
  UITS_3 ~~ UITS_4 + UITS_5 
  UITS_4 ~~ UITS_5 
  
  TEVR_1 ~~ TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  TEVR_2 ~~ TEVR_3 + TEVR_4 + TEVR_5
  TEVR_3 ~~ TEVR_4 + TEVR_5
  TEVR_4 ~~ TEVR_5
  
  BELA_1 ~~ BELA_2 + BELA_3 + BELA_4 + BELA_5
  BELA_2 ~~ BELA_3 + BELA_4 + BELA_5
  BELA_3 ~~ BELA_4 + BELA_5
  BELA_4 ~~ BELA_5
  
  VERA_1 ~~ VERA_2 + VERA_3 + VERA_4 + VERA_5
  VERA_2 ~~ VERA_3 + VERA_4 + VERA_5
  VERA_3 ~~ VERA_4 + VERA_5
  VERA_4 ~~ VERA_5
  
   # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit.sem.cu <- sem(model = sem.cu, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.sem.cu)
```

---

### Stationary SEM CU

```{r stationary-cu, eval = F}
stationary.cu <- '
  # Trait factor loadings
  T =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
       g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
       g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
       g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
       g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # Trait factor variance
  T ~~ T
  
  # State Residual factor loadings (constrained)
  SR1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  SR2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  SR3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  SR4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  SR5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # State Residual variances (constrained)
  SR1 ~~ vcs*SR1
  SR2 ~~ vcs*SR2
  SR3 ~~ vcs*SR3
  SR4 ~~ vcs*SR4
  SR5 ~~ vcs*SR5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5
  
  # Correlated measurement errors between the same methods over time 
  IDEA_1 ~~ cov1*IDEA_2 + cov1*IDEA_3 + cov1*IDEA_4 + cov1*IDEA_5
  IDEA_2 ~~ cov1*IDEA_3 + cov1*IDEA_4 + cov1*IDEA_5
  IDEA_3 ~~ cov1*IDEA_4 + cov1*IDEA_5
  IDEA_4 ~~ cov1*IDEA_5
  
  UITS_1 ~~ cov2*UITS_2 + cov2*UITS_3 + cov2*UITS_4 + cov2*UITS_5  
  UITS_2 ~~ cov2*UITS_3 + cov2*UITS_4 + cov2*UITS_5 
  UITS_3 ~~ cov2*UITS_4 + cov2*UITS_5 
  UITS_4 ~~ cov2*UITS_5 
  
  TEVR_1 ~~ cov3*TEVR_2 + cov3*TEVR_3 + cov3*TEVR_4 + cov3*TEVR_5 
  TEVR_2 ~~ cov3*TEVR_3 + cov3*TEVR_4 + cov3*TEVR_5
  TEVR_3 ~~ cov3*TEVR_4 + cov3*TEVR_5
  TEVR_4 ~~ cov3*TEVR_5
  
  BELA_1 ~~ cov4*BELA_2 + cov4*BELA_3 + cov4*BELA_4 + cov4*BELA_5
  BELA_2 ~~ cov4*BELA_3 + cov4*BELA_4 + cov4*BELA_5
  BELA_3 ~~ cov4*BELA_4 + cov4*BELA_5
  BELA_4 ~~ cov4*BELA_5
  
  VERA_1 ~~ cov5*VERA_2 + cov5*VERA_3 + cov5*VERA_4 + cov5*VERA_5
  VERA_2 ~~ cov5*VERA_3 + cov5*VERA_4 + cov5*VERA_5
  VERA_3 ~~ cov5*VERA_4 + cov5*VERA_5
  VERA_4 ~~ cov5*VERA_5
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit.stationary.cu <- sem(model = stationary.cu, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.stationary.cu)
```

---


### ML-SEM CU

From an ML-SEM perspective, there is no statistical difference between the CU model and the CUTS model. Therefore, we specify the ML-SEM CU model in the same way as the ML-SEM CUTS model. 

---

## Indicator Trait model {.tabset .tabset-fade}


### SEM IT 

```{r sem-it, eval = F}
sem.it <- '
  # Indicator-specific trait factor loadings
  IT_IDEA =~ IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5
  IT_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  IT_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  IT_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  IT_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5
  
  # Unique trait variances and covariances
  IT_IDEA ~~ IT_IDEA + IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_UITS ~~ IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_TEVR ~~ IT_TEVR + IT_BELA + IT_VERA
  IT_BELA ~~ IT_BELA + IT_VERA
  IT_VERA ~~ IT_VERA
  
  # Common state factor loadings
  CS1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  CS2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  CS3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  CS4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  CS5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  CS1 ~~ CS1
  CS2 ~~ CS2
  CS3 ~~ CS3
  CS4 ~~ CS4
  CS5 ~~ CS5
  
  # Unique state variances (included by default)
  
  # Means (included by default through argument)
'

fit.sem.it <- sem(model = sem.it, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.sem.it)
```

---

### Stationary SEM IT

```{r stationary-it, eval = F}
stationary.it <- '
  # Indicator-specific trait factor loadings
  IT_IDEA =~ 1*IDEA_1 + 1*IDEA_2 + 1*IDEA_3 + 1*IDEA_4 + 1*IDEA_5
  IT_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  IT_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  IT_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  IT_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5
  
  # Unique trait variances and covariances
  IT_IDEA ~~ IT_IDEA + IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_UITS ~~ IT_UITS + IT_TEVR + IT_BELA + IT_VERA
  IT_TEVR ~~ IT_TEVR + IT_BELA + IT_VERA
  IT_BELA ~~ IT_BELA + IT_VERA
  IT_VERA ~~ IT_VERA
  
  # Common state factor loadings (constrained)
  CS1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  CS2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  CS3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  CS4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  CS5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  CS1 ~~ vcs*CS1
  CS2 ~~ vcs*CS2
  CS3 ~~ vcs*CS3
  CS4 ~~ vcs*CS4
  CS5 ~~ vcs*CS5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit.stationary.it <- sem(model = stationary.it, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.stationary.it)
```

---

### ML-SEM IT 

```{r mlsem-it, eval = F}
mlsem.it <- '
  level: 1
      cs =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      IDEA ~~ IDEA + UITS + TEVR + BELA + VERA
      UITS ~~ UITS + TEVR + BELA + VERA
      TEVR ~~ TEVR + BELA + VERA
      BELA ~~ BELA + VERA
      VERA ~~ VERA
'

fit.mlsem.it <- sem(model = mlsem.it, data = dat.liss.long, cluster = "id")
summary(fit.mlsem.it)
```

---


## Reference Indicator model{.tabset .tabset-fade}

### SEM RI

```{r sem-RI, eval = F}
sem.RI <- '
  # General factor loadings
  G =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1 + 
       IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 +
       IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3 +
       IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 + 
       IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
        
  # General factor variance
  G ~~ G
  
  # Trait Residual factor loadings
  # IDEA serves as the reference indicator
  TR_UITS =~ UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5  
  TR_TEVR =~ TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 
  TR_BELA =~ BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5
  TR_VERA =~ VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5

  # Trait Residual factor variances
  TR_UITS ~~ TR_UITS
  TR_TEVR ~~ TR_TEVR
  TR_BELA ~~ TR_BELA
  TR_VERA ~~ TR_VERA
  
  # Allow covariances between trait residual factors
  TR_UITS ~~ TR_TEVR + TR_BELA + TR_VERA
  TR_TEVR ~~ TR_BELA + TR_VERA
  TR_BELA ~~ TR_VERA
  
  # Occasion factor loadings
  O1 =~ IDEA_1 + UITS_1 + TEVR_1 + BELA_1 + VERA_1  
  O2 =~ IDEA_2 + UITS_2 + TEVR_2 + BELA_2 + VERA_2 
  O3 =~ IDEA_3 + UITS_3 + TEVR_3 + BELA_3 + VERA_3
  O4 =~ IDEA_4 + UITS_4 + TEVR_4 + BELA_4 + VERA_4 
  O5 =~ IDEA_5 + UITS_5 + TEVR_5 + BELA_5 + VERA_5
  
  # Common state variances
  O1 ~~ O1
  O2 ~~ O2
  O3 ~~ O3
  O4 ~~ O4
  O5 ~~ O5
  
  # Unique state variances (included by default)
  
  # Means (included by default through argument)
'

fit.sem.RI <- sem(model = sem.RI, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.sem.RI)
```

---

### Stationary SEM RI

```{r stationary-RI, eval = F}
stationary.RI <- '
  # General factor loadings
  G =~ g1*IDEA_1 + g2*UITS_1 + g3*TEVR_1 + g4*BELA_1 + g5*VERA_1 + 
       g1*IDEA_2 + g2*UITS_2 + g3*TEVR_2 + g4*BELA_2 + g5*VERA_2 +
       g1*IDEA_3 + g2*UITS_3 + g3*TEVR_3 + g4*BELA_3 + g5*VERA_3 +
       g1*IDEA_4 + g2*UITS_4 + g3*TEVR_4 + g4*BELA_4 + g5*VERA_4 + 
       g1*IDEA_5 + g2*UITS_5 + g3*TEVR_5 + g4*BELA_5 + g5*VERA_5
        
  # General factor variance
  G ~~ G
  
  # Trait Residual factor loadings
  # IDEA serves as the reference indicator
  TR_UITS =~ 1*UITS_1 + 1*UITS_2 + 1*UITS_3 + 1*UITS_4 + 1*UITS_5  
  TR_TEVR =~ 1*TEVR_1 + 1*TEVR_2 + 1*TEVR_3 + 1*TEVR_4 + 1*TEVR_5 
  TR_BELA =~ 1*BELA_1 + 1*BELA_2 + 1*BELA_3 + 1*BELA_4 + 1*BELA_5
  TR_VERA =~ 1*VERA_1 + 1*VERA_2 + 1*VERA_3 + 1*VERA_4 + 1*VERA_5

  # Trait Residual factor variances
  TR_UITS ~~ TR_UITS
  TR_TEVR ~~ TR_TEVR
  TR_BELA ~~ TR_BELA
  TR_VERA ~~ TR_VERA
  
  # Allow covariances between trait residual factors
  TR_UITS ~~ TR_TEVR + TR_BELA + TR_VERA
  TR_TEVR ~~ TR_BELA + TR_VERA
  TR_BELA ~~ TR_VERA
  
  # Common state factor loadings (constrained)
  O1 =~ k1*IDEA_1 + k2*UITS_1 + k3*TEVR_1 + k4*BELA_1 + k5*VERA_1  
  O2 =~ k1*IDEA_2 + k2*UITS_2 + k3*TEVR_2 + k4*BELA_2 + k5*VERA_2 
  O3 =~ k1*IDEA_3 + k2*UITS_3 + k3*TEVR_3 + k4*BELA_3 + k5*VERA_3
  O4 =~ k1*IDEA_4 + k2*UITS_4 + k3*TEVR_4 + k4*BELA_4 + k5*VERA_4 
  O5 =~ k1*IDEA_5 + k2*UITS_5 + k3*TEVR_5 + k4*BELA_5 + k5*VERA_5
  
  # Common state variances (constrained)
  O1 ~~ vcs*O1
  O2 ~~ vcs*O2
  O3 ~~ vcs*O3
  O4 ~~ vcs*O4
  O5 ~~ vcs*O5
  
  # Unique state variances (constrained)
  IDEA_1 ~~ vus1*IDEA_1
  UITS_1 ~~ vus2*UITS_1
  TEVR_1 ~~ vus3*TEVR_1
  BELA_1 ~~ vus4*BELA_1
  VERA_1 ~~ vus5*VERA_1 
  
  IDEA_2 ~~ vus1*IDEA_2
  UITS_2 ~~ vus2*UITS_2
  TEVR_2 ~~ vus3*TEVR_2
  BELA_2 ~~ vus4*BELA_2
  VERA_2 ~~ vus5*VERA_2 
  
  IDEA_3 ~~ vus1*IDEA_3
  UITS_3 ~~ vus2*UITS_3
  TEVR_3 ~~ vus3*TEVR_3
  BELA_3 ~~ vus4*BELA_3
  VERA_3 ~~ vus5*VERA_3 
  
  IDEA_4 ~~ vus1*IDEA_4
  UITS_4 ~~ vus2*UITS_4
  TEVR_4 ~~ vus3*TEVR_4
  BELA_4 ~~ vus4*BELA_4
  VERA_4 ~~ vus5*VERA_4 
  
  IDEA_5 ~~ vus1*IDEA_5
  UITS_5 ~~ vus2*UITS_5
  TEVR_5 ~~ vus3*TEVR_5
  BELA_5 ~~ vus4*BELA_5
  VERA_5 ~~ vus5*VERA_5 
  
  # Means (constrained)
  IDEA_1 + IDEA_2 + IDEA_3 + IDEA_4 + IDEA_5 ~ m1*1
  UITS_1 + UITS_2 + UITS_3 + UITS_4 + UITS_5 ~ m2*1
  TEVR_1 + TEVR_2 + TEVR_3 + TEVR_4 + TEVR_5 ~ m3*1
  BELA_1 + BELA_2 + BELA_3 + BELA_4 + BELA_5 ~ m4*1
  VERA_1 + VERA_2 + VERA_3 + VERA_4 + VERA_5 ~ m5*1 
'

fit.stationary.RI <- sem(model = stationary.RI, data = dat.liss, missing = 'ML', meanstructure = T, orthogonal = T)
summary(fit.stationary.RI)
```

---


### ML-SEM RI 


```{r mlsem-RI, eval = F}
mlsem.RI <- '
  level: 1
      O =~ IDEA + UITS + TEVR + BELA + VERA
  level: 2
      G =~ IDEA + UITS + TEVR + BELA + VERA
      
      IDEA ~~ 0*IDEA
      UITS ~~ UITS + TEVR + BELA + VERA
      TEVR ~~ TEVR + BELA + VERA
      BELA ~~ BELA + VERA
      VERA ~~ VERA
'

fit.mlsem.RI <- sem(model = mlsem.RI, data = dat.liss.long, cluster = "id")
summary(fit.mlsem.RI)
```

---
