---
title: "Data cleaning and preparation (R)"
---

---

The raw LISS data was cleaned and prepared before the models were fitted. Below you can find the *R* code that was used. The most important steps are:

- selection of variables and waves of data,
- random selection of a single individual from each household,
- renaming the variables in the data set, 
- inspecting missingness patterns and deleting individuals with only missing values, and
- restructuring the data into long-format. 


```{r setup, message=F, warning=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(data.table)
library(Hmisc) # Import SPSS files
library(visdat) # Visualize missing data
library(sjlabelled) # Working with labeled data (often imported from SPSS)
library(panelr) # Tools for wrangling panel data
library(purrr)
```

```{r data, eval=F}
# Import sample characteristics files
set.seed(12345) # Seed for random selecting later in pipe
dat_liss <- map(c("./data/LISS/Wave 1/cp08a_1p_EN.sav", 
                  "./data/LISS/Wave 2/cp09b_1.0p_EN.sav",
                  "./data/LISS/Wave 3/cp10c_1.0p_EN.sav",
                  "./data/LISS/Wave 4/cp11d_1.0p_EN.sav",
                  "./data/LISS/Wave 5/cp12e_1.0p_EN.sav",
                  "./data/LISS/Wave 6/cp13f_EN_1.0p.sav",
                  "./data/LISS/Wave 7/cp14g_EN_1.0p.sav",
                  "./data/LISS/Wave 8/cp15h_EN_1.0p.sav"),
                spss.get, use.value.labels = F, allow = c("_")) %>% 
  reduce(full_join, by = "nomem_encr") %>%
  group_by(nohouse_encr.x) %>%
  sample_n(1) %>% # Randomly select a single individual from each household
  ungroup() 

# Select Diener's Satisfaction with Life Scale from first 5 waves of data
dat_LS <- select(dat_liss, 
                 nomem_encr,
                 num_range(prefix = "cp08a0", range = 14:18), 
                 num_range(prefix = "cp09b0", range = 14:18), 
                 num_range(prefix = "cp10c0", range = 14:18), 
                 num_range(prefix = "cp11d0", range = 14:18), 
                 num_range(prefix = "cp12e0", range = 14:18))
  
# Explore missingness in the data
vis_miss(select(dat_LS, ends_with("014"))) # Select question 014 for missingness 

# Data characteristics I
W <- 5 # Number of waves
M <- 5 # Number of indicators

# Create column names
item_names <- c("IDEA", "UITS", "TEVR", "BELA", "VERA")
waves <- paste0("_", 1:W)
x <- expand.grid(item_names, waves)
var_names <- paste0(x[,1], x[,2])

## Rename columns
names(dat_LS) <- c("person", var_names)
dat_LS <- remove_all_labels(dat_LS) # Remove labels
dat_LS <- filter_at(dat_LS, vars(IDEA_1:VERA_5), any_vars(!is.na(.))) # Remove rows with all missing. 

# Data characteristics II
n <- nrow(dat_LS)

## Create long format 
dat_LS_long <- long_panel(dat_LS, prefix = "_", begin = 1, end = 5, label_location = "end")

# Export to Mplus
write_dat <- partial(fwrite, sep = "\t", col.names = F, row.names = F, na = ".") # Create partial function for quick use
write_dat(dat_LS, file = "./Mplus - Applied factor models/LISS_LS_wide.dat")
write_dat(dat_LS_long, file = "./Mplus - Applied factor models/LISS_LS_long.dat")
```

